### 2019-03-28 rebuttal.txt
Starting the rebuttal.

We thank the reviewers for their thorough reading of the article and 
for their valuable feedback.  We shall apply the reviewers' comments
to improve the article for the final version.

On the reviewers' main concerns:

R1: TODO

R2: TODO

R3 notes that we do not define a dependency pair notion or any automation.

These two points are currently named as lines for future work.  We believe
that it would not benefit the present work to do too much at once.
Automation is challenging because it requires the development of several new
ideas; setting up a dependency pair approach would be a full, mostly separate
line of research.  (However, it is worth noting that a future definition of
dependency pairs is likely to benefit greatly from the existence of a
reduction pair, like the one we define here).